---
title: "project-2"
output: pdf_document
date: "2025-11-19"
---

## **Task 1**
```{r task 1}
medicaldata <- read.csv("medicaldata2.csv")

medicaldata$Gallstone_Status <- as.factor(medicaldata$Gallstone_Status)
medicaldata$Comorbidity <- factor(medicaldata$Comorbidity, ordered = TRUE, levels = c(0, 1, 2, 3), labels = c("No comorbidities present", "One comorbid condition", "Two comorbid conditions", "Three of more comorbid conditions"))
medicaldata$CAD <- as.factor(medicaldata$CAD)
medicaldata$Diabetes_Mellitus <- as.factor(medicaldata$Diabetes_Mellitus)
medicaldata$Hepatic_Fat_Accumulation <- factor(medicaldata$Hepatic_Fat_Accumulation, ordered = TRUE, levels = c(0, 1, 2, 3), labels = c("No fat accumulation", "Grade 1 (mild)", "Grade 2 (moderate)", "Grade 3 (severere)"))
medicaldata$Age_Class <- factor(medicaldata$Age_Class, ordered = TRUE, levels = c("Younger", "Lower Middle", "Upper Middle", "Elder"))
```

## **Task 2**
```{r task 2}
plot(medicaldata$Body_Protein_Content, medicaldata$Total_Body_Fat_Ratio, xlab = "Body Protein Content (%)", ylab = "Total Body Fat Ratio (%)", main = "Total Body Fat Ratio versus Body Protein Content")

task2model <- lm(Total_Body_Fat_Ratio ~ Body_Protein_Content, data = medicaldata)

plot(task2model$fitted.values, task2model$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals versus Fitted Values")
abline(h = 0)

qqnorm(task2model$residuals, main = "Normality Q-Q Plot")
qqline(task2model$residuals)

shapiro.test(task2model$residuals)

pearson_coefficient <- cor(medicaldata$Body_Protein_Content, medicaldata$Total_Body_Fat_Ratio, method = "pearson")
pearson_coefficient

task2model$coefficients
summary(task2model)
```

H0: Beta_1 = 0
H1: Beta_1 != 0
alpha = 0.05

F = 307.7
p < 2.2e-16

Decision: Reject H0. p < 2.2e-16 < 0.05 = alpha
Conclusion: There is evidence of linear association between Body Protein Content and Body Fat Ratio.

Regression Line Equation: Total_Body_Fat_Ratio = -2.624164 * Body_Protein_Content + 70.205708
Pearson Correlation Coefficient: -0.7029279
R^2: 0.4941

The linearity and equal variance assumptions are met. The residuals appear randomly distributed around the value zero in the Residuals versus Fitted Values plot. The normality assumption is also met. The p-value obtained from the Shapiro-Wilk test is p = 0.5171 > 0.05. The Pearson Correlation Coefficient r = -0.7029279 indicates a strong negative linear association between Total Body Fat Ratio and Body Protein Content. The R^2 value R^2 = 0.4941 means that 49.41% of the variability in Total Body Fat Ratio is explained by Body Protein Content.

## **Task 3**
```{r task 3}
task3model <- lm(Total_Body_Water ~ Height + Weight + Body_Protein_Content, data = medicaldata)

plot(task3model$fitted.values, task2model$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals versus Fitted Values")
abline(h = 0)

qqnorm(task3model$residuals, main = "Normality Q-Q Plot")
qqline(task3model$residuals)

shapiro.test(task3model$residuals)

library(lmtest)
dwtest(task3model)

summary(task3model)
```
H0: Beta_Height = Beta_Weight = Beta_Body_Protein_Content = 0
H1: At least one Beta != 0

F = 368.6
p < 2.2e-16

Decision: Reject H0. p < 2.2e-16 < 0.03 = alpha

Conclusion: There is enough evidence to conclude that Height, Weight, and Body Protein Content together significantly explain variation in Total Body Water.


The linearity and equal variance assumptions are met. The residuals appear randomly distributed around the value zero in the Residuals versus Fitted Values plot. The normality assumption is not met. The p-value obtained from the Shapiro-Wilk test is p = 2e-12 < 0.03. The independence assumption is met. The p-value obtained from the Durbin-Watson test is p = 0.4822 > 0.03. The value of R^2 is R^2 = 0.7794 and the value of R^2_adj is R^2_adj = 0.7773. We can reasonably conclude that the combination of Height, Weight, and Body Protein Content together significantly explains variation in Total Body Water. Because of this, it is also important to test for the individual variables to determine the individual effects on Total Body Water.


Test 1: Height

H0: Beta_Height = 0
H1: Beta_Height != 0

t = 12.055
p < 2e-16

Decision: Reject H0. p < 2e-16 < 0.03 = alpha

Conclusion: There is enough evidence to conclude that Height significantly explains variation in Total Body Water.


Test 2: Weight

H0: Beta_Weight = 0
H1: Beta_Weight != 0

t = 16.307
p < 2e-16

Decision: Reject H0. p < 2e-16 < 0.03 = alpha

Conclusion: There is enough evidence to conclude that Weight significantly explains variation is Total Body Water.


Test 3: Body Protein Content

H0: Beta_Body_Protein_Content = 0
H1: Beta_Body_Protein_Content != 0

t = 0.927
p = 0.355

Decision: Do not reject H0. p = 0.355 > 0.03 = alpha

Conclusion: There is not enough evidence to conclude that Body Protein Content significantly explains variation in Total Body Water.


Based on the individual tests, we can conclude that Height and Weight are statistically significant predictors for Total Body Water. Body Protein Content is not a statistically significant predictor. The R^2 and R^2 adjusted values indicate that approximately 78% of the variability in Total Body Water can be explained by Height, Weight, and Body Protein Content.

## **Task 4**
```{r task 4}
```

## **Task 5**
```{r task 5}
```